Getting rid of the \texttt{console.log}:\\Debugging of RxJS-based Applications

### 2020-11-08 slides.tex
Future Work

\note[itemize] {
    \item Frustration -> Reason Research topic
    \item There are concepts and solutions for RP debugging
    \item Reactive Debugger, Scala \cite{10.1145/2884781.2884815}
    \item Improved logging tools rxjs-spy
    \item Visualizers rxviz, rxjs visualizer
    \item But why does nobody seem to use them?
    \item Why are they not popular?
    \item -> Research Questions
}

@@

\begin{frame}{Quotes}
    \begin{vfilleditems}
		\item \textit{``[Reactive Programming is] a good way for \textbf{composing} multiple data sources.''}
		\item \textit{``In \textbf{99\%} of all cases, I add \textbf{console.log} statements [\dots] trying to understand what's happening.''}
	\end{vfilleditems}
\end{frame}

@@

\note[itemize] {
    \item Foo
}

@@

\begin{frame}[fragile]{Key Take Aways}
    \begin{columns}[T]
        \begin{column}{.6\linewidth}
            \vspace{3em}
            \begin{itemize}
                \item The ``\textbf{to hammer a screw}''-problem\bigskip
                \item Tendency to \textbf{manual} code \textbf{modifications}\bigskip
                \item \textbf{Pessimistic} opinions on debugging of RxJS-based applications
	        \end{itemize}
        \end{column}
        \begin{column}{.4\linewidth}
            \begin{figure}
                \centering
                \includegraphics[width=.7\linewidth]{figures/hammer-screw.jpeg}
                \caption{\tiny{Source: Manuel Alabor}}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\note[itemize] {
    \item ... todo
    \item Based on this
    \item Observational study to validate data
    \item Using the following hypothesis:
}

@@

\begin{frame}[fragile]{Hypothesis}
    \begin{quote}
        If software engineers must solve an RxJS-based problem, then they \textbf{will instrument the code manually} in order to understand its behavior.
    \end{quote}
\end{frame}

\note[itemize] {
    \item Foo
}

@@

\note[itemize] {
    \item All modified/augmented source code manually
    \item -> Trace logs
    \item Two tried traditional debugging
    \item No one extracted code
    \item No one used additional libraries
    \item On told afterwards: Would have been their next step
    \item Based on this evidence: Hypothesis is correct
}

\begin{frame}[fragile]{Interpretation}
    \begin{vfilleditems}
        \item \textbf{Four} professional software engineers
        \item Two currently \textbf{work} with RxJS
        \item Three have \textbf{two or more years} of \textbf{experience} with RxJS
        \item One of them uses RxJS to develop \textbf{backend applications}
    \end{vfilleditems}
\end{frame}

\note[itemize] {
    \item All used manual trace logs
    \item Two tried traditional debugging
    \item No one extracted code
    \item No one used additional libraries
    \item On told afterwards: Would have been their next step
}

\section{Conclusion}
\begin{frame}{}

\end{frame}

\section{Future Work}
\begin{frame}[fragile]{Future Work}
    \begin{vfilleditems}
        \item Manuel Alabor \cite{10.1145/2501654.2501666}
        \item Manuel Alabor
        \item Manuel Alabor
    \end{vfilleditems}
\end{frame}

\begin{frame}[plain, noframenumbering]
  \centering
  \vfill
  {\fontsize{40}{50}\selectfont Q \& A}
  \vfill
\end{frame}

>>>>

\note[itemize] {
    \item Frustration -> Reason Research topic
    \item There are good concepts and solutions for RP debugging
    \item Reactive Debugger, Scala \cite{10.1145/2884781.2884815}
    \item Improved logging tools rxjs-spy \cite{rxjsspy}
    \item Visualizers rxviz \cite{rxviz}
    \item Visualizers rxfiddle \cite{10.1145/3180155.3180156}
    \item But why does nobody seem to use them?
    \item Why unpopular?
    \item -> Research Questions
}

@@

\begin{frame}{Quotes}
    \begin{vfilleditems}
		\item[\Large{``}] \textit{[Reactive Programming is] a good way for \textbf{composing} multiple data sources.}
		\item[\Large{``}] \textit{In \textbf{99\%} of all cases, I add \textbf{console.log} statements [\dots] trying to understand what's happening.}
	\end{vfilleditems}
\end{frame}

@@

\note[itemize] {
    \item Again: Common practice to modify code
    \item Not only trace logs
    \item Extraction of code parts to extraneous tools
    \item Visualizers rx-viz \cite{rxviz}
    \item Code sandboxes StackBlitz
    \item Try to use debugger
    \item Debuggers often not that useful
}

@@

\begin{frame}[fragile]{Key Take Aways}
    \begin{columns}[T]
        \begin{column}{.6\linewidth}
            \vspace{3em}
            \begin{itemize}
                \item The ``\textbf{to hammer a screw}''-problem\bigskip
                \item Tendency to \textbf{manual} code \textbf{modifications}\bigskip
                \item Rather \textbf{negative} experience with debugging RxJS-based applications
	        \end{itemize}
        \end{column}
        \begin{column}{.4\linewidth}
            \begin{figure}
                \centering
                \includegraphics[width=.7\linewidth]{figures/hammer-screw.jpeg}
                \caption{\tiny{Source: Manuel Alabor}}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\note[itemize] {
    \item Traditional debugger feels wrong
    \item Ask a question which they cannot answer (hammer/screw) -> Stack trace
    \item Fallback to manual code modification
    \item Trace log statements
    \item Extract pieces to sandboxes
    \item Overall: Pessimistic view on debugging RxJS RP
    \item Based on this:
    \item Study to validate this...
    \item Hypothesis coming up:
}

@@

\begin{frame}[fragile]{Hypothesis}
    \begin{vfilleditems}
        \item[\Large{``}] If software engineers must solve an RxJS-based problem, then they \textbf{will instrument the code manually} in order to understand its behavior.
    \end{vfilleditems}
\end{frame}

\note[itemize] {
    \item Manual intervention repeating theme
    \item Both Interviews and War Stories
    \item Thats why focus on this in validation study
    \item Dont care what kind (logs vs extraction)
}

@@

\note[itemize] {
    \item All modified/augmented source code manually
    \item -> Trace logs
    \item Two tried traditional debugging
    \item No one extracted code
    \item No one used additional tools
    \item One told afterwards: Would have been their next step
    \item Based on experiment data: Hypothesis is correct
}

\section{Conclusion}

\begin{frame}[fragile]{Interpretation}
    \begin{vfilleditems}
        \item Engineers \textbf{know} RP debugging tools, but \textbf{do not use} them
        \item Engineers use \textbf{imperative}, but actually require ``\textbf{reactive}'' debuggers
        \item \textbf{Manual intervention} is the ``\textbf{last resort}''
    \end{vfilleditems}
\end{frame}

\note[itemize] {
    \item Interviews + War Stories + Study = ?
    \item 7 of 14 peers know specific RxJS debugging tools
    \item But they dont use them
    \item Instead: Traditional tools \textbf{easily accessible}
    \item Final way to go:
    \item Trace logs and code extraction
}

\begin{frame}[fragile]{Answer}
    \begin{enumerate}
        \item[\color{gray}{RQ1}] \color{gray}{What challenges do software engineers face when debugging RxJS-based applications?}
        \vfill\item[\Large{``}] \color{white}{The most significant challenge software engineers face [\dots] is to know \textbf{when} they should apply \textbf{what tool} to resolve their current problem in the \textbf{most efficient} way.}
    \end{enumerate}
\end{frame}

\note[itemize] {
    \item Challenge: What tool, when and how
    \item Often: The "next best" as we saw: Traditional tools
    \item Often: Modify code
    \item Less often: Specific tools
    \item What can we do about it?
    \item Answer question 2
}

\begin{frame}[fragile]{Answer}
    \begin{enumerate}
        \item[\color{gray}{RQ2}] \color{gray}{How can the experience of software engineers during the debugging process of RxJS-based applications be improved?}
        \vfill\item[\Large{``}] \color{white}{[\dots] \textbf{improve} the \textbf{experience} of debugging [\dots] by providing RP specific debugging utilities where software engineers expect them the most: \textbf{Fully integrated} [\dots] their IDE [\dots]}
    \end{enumerate}
\end{frame}

\note[itemize] {
    \item Improve the dev user experience
    \item Lower the boundary to use the right tool
    \item At the right time
    \item Integrate the right tools where they are expected the most
    \item Example: Async Stack Traces in Google Chrome \cite{chromeasync}
}

\section{Future Work}

\begin{frame}[fragile]{Future Work}
    \begin{vfilleditems}
        \item \textbf{Propose} and \textbf{implement} solution to \textbf{prevent} manual code \textbf{modification}
        \item \textbf{Evaluate} proposal on \textbdf{effectiveness} to answer RQ3
    \end{vfilleditems}
\end{frame}

\note[itemize] {
    \item Wraps up our plans for the future
    \item Work on a concrete prototype
    \item Find solution for trace log problem
    \item Without manual intervention
    \item Validate effectiveness and answer RQ3
}

\begin{frame}[fragile]{Sneak Preview}
    \begin{figure}[H]
        \centering
        \includegraphics[height=0.6\textheight]{figures/tweet-rxjs-debugging-extension.png}
        \caption{\tiny{Source: \url{https://twitter.com/swissmanu/status/1313412280072232960}}}
    \end{figure}
\end{frame}

\note[itemize] {
    \item Small sneak preview
    \item Things are progressing
    \item Thank you for your time and attention!
}

\begin{frame}[plain, noframenumbering]
  \centering
  \vfill
  {\fontsize{40}{50}\selectfont\color{white}{Q \& A}}
  \vfill
\end{frame}

\note[itemize] {
    \item Happy for input
    \item Happy to answer questions
}
